---
title: "tidytuesday David Robinson - transit Analysis"
output: 
  html_notebook:
    highlight: tango
    df_print: paged
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
    number_sections: yes
    toc_depth: 6
  html_document:
    toc: yes
    toc_depth: '6'
    df_print: paged
---

# source

https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-01-05/readme.md

https://github.com/dgrtwo/data-screencasts/blob/master/2021_01_05_transit_costs.Rmd

https://www.youtube.com/watch?v=8jNQzce13SE

# Options & Settings

CSS for scrollable output & Header colors 

```{css, echo=FALSE}
.scroll-100 {
  max-height: 100px;
  overflow-y: auto;
  background-color: inherit;
}
```


```{r}
library(tidyverse)
library(tidytuesdayR)
library(scales)
library(glue)
```


```{r}
theme_set(theme_minimal())
```


# Data 

Download the weekly data

```{r}
tt <- tidytuesdayR::tt_load("2021-01-05")

head(tt)
```

# Summary

```{r}
str(tt)
```

```{r}
library(countrycode)
```


```{r}
transit_cost <- tt$transit_cost %>% 
  mutate_at(vars(country,city,line), as.factor) %>% 
  mutate_at(vars(start_year,end_year, real_cost), as.numeric)

str(transit_cost)
```

```{r}
summary(transit_cost)
```


```{r }
transit_cost %>% 
  add_count(city, name = "city_counts") %>% 
  select(city, city_counts) %>% 
  unique() %>% 
  arrange(desc(city_counts))
```

## Freq. function

```{r}
top_freq <- function(df, topn=10){

  df %>% 
    select_if(is.factor) %>% 
    imap(function(feature_value, feature_name) {
      
      count(data.frame(x = feature_value), x) %>% 
        filter(!is.na(x)) %>%
        slice_max(n = topn, order_by = n) %>% 
      
          ggplot(aes(x = fct_reorder(x, n), y = n)) +
          geom_col() +
          coord_flip() +
          theme(axis.text.x = element_text(angle=90)) +
          labs(title = feature_name,
               x = feature_name)
  
      })

}

top_freq(transit_cost, 15)
```


```{r}
transit_cost <- transit_cost %>% 
  filter(!is.na(e)) %>% 
  mutate(country = as.character(country),
         
         # if you don't convert to "char" above then due to factors it will return NA in country
         country_code = ifelse(country == "UK", "GB", country),
         country = countrycode(country_code, "iso2c", "country.name"),
         country = as.factor(country),
         tunnel_per = tunnel / length,
         rr = ifelse(rr, "Railroad", "Not Railroad"))

transit_cost
```

```{r}
transit_cost %>% 
  count(city, country, sort = TRUE)
```

```{r}
top_freq(transit_cost, 20)
```

```{r}
transit_cost %>%
  add_count(line, sort = TRUE) %>% 
  select(line, n) %>% 
  unique() %>%
  slice_max(order_by = n, n =15) %>% 
  pull(line)
```


```{r}
 transit_cost %>%
  add_count(line, sort = TRUE) %>% 
  filter(line %in% (transit_cost %>%
                    add_count(line, sort = TRUE) %>% 
                    select(line, n) %>% 
                    unique() %>%
                    slice_max(order_by = n, n =15) %>% 
                    pull(line))
         ) %>%
  
  ggplot(aes(x = fct_reorder(line, n), y = n, fill = year)) +
  geom_col() +
  coord_flip() +
  theme(axis.text.x = element_text(angle=90)) +
  labs(title = "top 15 lines segmented by years as color ",
       x = "line",
       fill = "year")
```


```{r}
 transit_cost %>%
  add_count(line, sort = TRUE) %>% 
  filter(line %in% (transit_cost %>%
                    add_count(line, sort = TRUE) %>% 
                    select(line, n) %>% 
                    unique() %>%
                    slice_max(order_by = n, n =15) %>% 
                    pull(line))
         ) %>%
  
  ggplot(aes(x = fct_reorder(line, n), y = n, fill = year, group = -year)) +
  geom_col() +
  coord_flip() +
  theme(axis.text.x = element_text(angle=90)) +
  labs(title = "top 15 lines segmented by years as color ",
       x = "line",
       fill = "year") +
  scale_fill_viridis_b()
```



```{r}
 transit_cost %>%
  add_count(line, sort = TRUE) %>% 
  filter(line %in% (transit_cost %>%
                    add_count(line, sort = TRUE) %>% 
                    select(line, n) %>% 
                    unique() %>%
                    slice_max(order_by = n, n =15) %>% 
                    pull(line))
         ) %>%
  mutate(line = fct_reorder(line, year)) %>% 
  ggplot(aes(xmin = start_year, xmax = end_year, y = line, color = country),
         size = real_cost) +
  geom_errorbarh(height = .2) +
  labs(title = "Top 15 lines wrt year",
       x = "Year",
       y = "",
       color = "Country")
```

Above chart shows that lines of same name appears in multiple countries, which is not the right way of creating this chart

## reorder_within()

```{r fig.width=8, fig.height=8}
 transit_cost %>%
  add_count(line, sort = TRUE) %>% 
  filter(line %in% (transit_cost %>%
                    add_count(line, sort = TRUE) %>% 
                    select(line, n) %>% 
                    unique() %>%
                    slice_max(order_by = n, n =15) %>% 
                    pull(line))
         ) %>%
  
  # using reorder_within() instead of fct_reorder()
  mutate(line = reorder_within(line, year, country)) %>% 
  ggplot(aes(xmin = start_year, xmax = end_year, y = line, color = country),
         size = real_cost) +
  geom_errorbarh(height = .8) +
  labs(title = "Top 15 'Project lines' wrt year now segrgated country wise",
       subtitle = "Top 15 line names turned into more when segregated using reorder_within",
       x = "Year",
       y = "",
       color = "Country")
```


```{r fig.width=8, fig.height=8}
 transit_cost %>%
  add_count(line, sort = TRUE) %>% 
  filter(line %in% (transit_cost %>%
                    add_count(line, sort = TRUE) %>% 
                    select(line, n) %>% 
                    unique() %>%
                    slice_max(order_by = n, n =15) %>% 
                    pull(line))
         ) %>%
  
  # using reorder_within() instead of fct_reorder()
  mutate(line = reorder_within(line, year, country)) %>% 
  ggplot(aes(xmin = start_year, xmax = end_year, y = line, color = country),
         size = real_cost) +
  geom_errorbarh(height = .8) +
  labs(title = "Top 15 'Project lines' wrt year now segrgated country wise",
       subtitle = "Top 15 line names turned into more when segregated using reorder_within",
       x = "Year",
       y = "",
       color = "Country") +
  scale_y_reordered()
```


```{r}
 transit_cost %>%
  add_count(line, sort = TRUE) %>% 
  filter(line %in% (transit_cost %>%
                      group_by(country, line) %>% 
                      add_count(line, sort = TRUE) %>% 
                      select(line, n) %>% 
                      unique() %>% 
                      ungroup() %>% 
                      slice_max(order_by = n, n =15) %>% 
                      pull(line))
         ) %>%
  mutate(line = fct_reorder(line, year)) %>% 
  ggplot(aes(xmin = start_year, xmax = end_year, y = line, color = country),
         size = real_cost) +
  geom_errorbarh(height = .2) +
  labs(title = "Top 15 grouped by country & line ",
       x = "Year",
       y = "",
       color = "Country")
```

```{r}
 transit_cost %>%
  add_count(line, sort = TRUE) %>% 
  filter(line %in% (transit_cost %>%
                      group_by(country, line) %>% 
                      add_count(line, sort = TRUE) %>% 
                      select(line, n) %>% 
                      unique() %>% 
                      ungroup() %>% 
                      slice_max(order_by = n, n =15) %>% 
                      pull(line))
         ) %>%
  mutate(line = fct_reorder(line, year)) %>% 
  ggplot(aes(xmin = start_year, xmax = end_year, y = line, color = city),
         size = real_cost) +
  geom_errorbarh(height = .2) +
  labs(title = "Top 15 grouped by country & line ",
       x = "Year",
       y = "",
       color = "City")
```


```{r}
transit_cost %>% 
  filter(country == "United States") %>% 
  mutate(line = fct_reorder(line, year)) %>% 
  ggplot(aes(xmin = start_year, xmax = end_year, y = line),
         color = city,
         size = real_cost) +
  geom_errorbarh(height = .2) +
  labs(x = "Year",
       y = "",
       color = "City")
```

```{r}
transit_cost %>% 
  ggplot(aes(cost_km_millions)) +
  geom_histogram() +
  scale_x_continuous(labels = dollar) +
  labs(title = "Distribution of Cost of constructing lines",
       x = "Cost / KM (USD Millions)")
```

## Cost/km {.tabset .tabset-fade .tabset-pills}

### Data prep.

```{r}
transit_cost %>% 
  filter(!is.na(cost_km_millions),
         tunnel_per == 1) %>% 
  mutate(country = fct_lump(country, 10))
```

### Top Countries box plot 

```{r}
transit_cost %>% 
  filter(!is.na(cost_km_millions),
         tunnel_per == 1) %>% 
  mutate(country = fct_lump(country, 10)) %>% 
  add_count(country) %>% 
  mutate(country = glue("{country} ({n})"),
         country = fct_reorder(country, cost_km_millions)) %>% 
  
  ggplot(aes(cost_km_millions, country)) +
  geom_boxplot() +
  scale_x_continuous(labels = dollar) +
  labs(x = "Cost / km (Mn USD)", y = "")
  
```

### +colored

```{r}
transit_cost %>% 
  filter(!is.na(cost_km_millions),
         tunnel_per == 1) %>% 
  mutate(country = fct_lump(country, 10)) %>% 
  add_count(country) %>% 
  mutate(country = glue("{country} ({n})"),
         country = fct_reorder(country, cost_km_millions)) %>% 
  
  ggplot(aes(cost_km_millions, country, color = country)) +
  geom_boxplot() +
  geom_jitter() +
  scale_x_continuous(labels = dollar) +
  labs(x = "Cost / km (Mn USD)", y = "")
  
```

### +India

```{r}
transit_cost %>% 
  filter(!is.na(cost_km_millions),
         tunnel_per == 1,
         country == "India") %>% 
  mutate(city = fct_lump(city, 10)) %>% 
  add_count(city) %>% 
  mutate(city = glue("{city} ({n})"),
         city = fct_reorder(city, cost_km_millions)) %>% 
  
  ggplot(aes(cost_km_millions, city)) +
  geom_boxplot() +
  scale_x_continuous(labels = dollar) +
  labs(x = "Cost/km (Millions USD)",
       y = "") 
```

### +India(without tunner_per=1)

```{r}
transit_cost %>% 
  filter(!is.na(cost_km_millions),
         # tunnel_per == 1,
         country == "India") %>% 
  mutate(city = fct_lump(city, 10)) %>% 
  add_count(city) %>% 
  mutate(city = glue("{city} ({n})"),
         city = fct_reorder(city, cost_km_millions)) %>% 
  
  ggplot(aes(cost_km_millions, city)) +
  geom_boxplot() +
  scale_x_continuous(labels = dollar) +
  labs(x = "Cost (Millions USD)",
       y = "") 
```

### +China

```{r}
transit_cost %>% 
  filter(!is.na(cost_km_millions),
         tunnel_per == 1,
         country == "China") %>% 
  mutate(city = fct_lump(city, 10)) %>% 
  add_count(city) %>% 
  mutate(city = glue("{city} ({n})"),
         city = fct_reorder(city, cost_km_millions)) %>% 
  
  ggplot(aes(cost_km_millions, city)) +
  geom_boxplot() +
  scale_x_continuous(labels = dollar) +
  labs(x = "Cost/km (Millions USD)",
       y = "") 
```

## errorbarh {.tabset .tabset-fade .tabset-pills}

### India 

```{r}
transit_cost %>%
  filter(country == "India",
         !is.na(start_year),
         !is.na(end_year)) %>%
  mutate(city = fct_lump(city, 5)) %>%
  mutate(line = fct_reorder(line, year)) %>%
  ggplot(aes(xmin = start_year, xmax = end_year, y = line,
             color = city,
             size = real_cost)) +
  geom_errorbarh(height = .1) +
  labs(title = "India line Projects",
       x = "Year",
       y = "",
       color = "City")
```


### China 

```{r}
transit_cost %>% 
  filter(country == "China",
         city == "Shanghai",
         !is.na(start_year),
         !is.na(end_year)) %>% 
  mutate(city = fct_lump(city, 5)) %>% 
  mutate(line = fct_reorder(line, year)) %>% 
  
  ggplot(aes(xmin = start_year, xmax = end_year, y = line, color = city, size = real_cost)) +
  geom_errorbarh(height = .2) +
  labs(title = "China line projects",
       x = "Year", y ="", color = "City"
       )
```

## India city wise {.tabset .tabset-fade .tabset-pills}

```{r}
transit_cost %>% 
  filter(country == "India")
```

### Projects timeline

```{r}
transit_cost %>% 
  filter(country == "India",
         !is.na(start_year),
         !is.na(end_year)) %>%
  mutate(city = fct_reorder(city, year, .fun = max)) %>% 
  
  ggplot(aes(xmin = start_year, xmax = end_year, y = city, color = rr, size = tunnel_per)) +
  geom_errorbarh(height = 0.2) +
  labs(title = "Indian cities Projects timeline",
       subtitle = "Size based on tunner_per")
  
```

### Length in Projects

```{r}
transit_cost %>% 
  filter(country == "India",
         # !is.na(start_year),
         # !is.na(end_year)
         ) %>%
  mutate(city = fct_reorder(city, length, sum)) %>% 
  
  ggplot(aes(x = length, y = city, fill = line)) +
  geom_col() +
  labs(title = "Total Lenth of Projects across Indian cities",
       subtitle = "color based on Project Lines")
  
```

### Projects cost

```{r}
transit_cost %>% 
  filter(country == "India") %>%
  mutate(city = fct_reorder(city, real_cost, sum)) %>% 
  
  ggplot(aes(x = real_cost, y = city, fill = year, group = -year)) +
  geom_col() +
  scale_x_continuous(label = scales::comma_format()) +
  labs(title = "Total real cost of Projects across Indian cities",
       subtitle = "color based on Year of Project Lines",
       fill = "Years") +
  scale_fill_viridis_b()
```

## Median Cost

```{r}
transit_cost %>% 
  filter(tunnel_per == 1,
         end_year <= 2020,
         country == "China") %>% 
  group_by(year = (year %/% 5) * 5) %>% 
  summarise(median_cost_km = median(cost_km_millions),
            n = n()) %>% 
  
  ggplot(aes(year, median_cost_km)) +
  geom_line() +
  geom_point(aes(size = n))
```

```{r}
transit_cost %>% 
  filter(
    # tunnel_per == 1,
         end_year <= 2020,
         country == "India") %>% 
  group_by(year = (year %/% 5) * 5) %>% 
  summarise(median_cost_km = median(cost_km_millions),
            n = n()) %>% 
  
  ggplot(aes(year, median_cost_km)) +
  geom_line() +
  geom_point(aes(size = n))
```


```{r}
transit_cost %>% 
  filter(
    # tunnel_per == 1,
         # end_year <= 2020,
         country == "India") %>% 
  group_by(city, 
           year = (year %/% 5) * 5
           ) %>% 
  summarise(median_cost_km = median(cost_km_millions),
            n = n()) %>% 
  
  ggplot(aes(year, median_cost_km, color = city)) +
  geom_line() +
  geom_point(aes(size = n))
```

```{r}
transit_cost %>%
  filter(
    # tunnel_per == 1,
    # end_year <= 2020,
         country == "India") %>%
  mutate(year = (year %/% 5) * 5,
         city = fct_lump(city, 5)) %>% 
  
  ggplot(aes(year, cost_km_millions, group = year)) +
  geom_boxplot(outlier.size = -1) +
  geom_jitter(aes(color = city), height = 0, width = 1) +
  expand_limits(y = 0) +
  labs(y = "Cost/km (Real USD in Millions)",
       x = "Year",
       title = "Cost distr./ km in China")
```

```{r}
library(ggforce)
```

```{r}
transit_cost %>%
  filter(
    # tunnel_per == 1,
    # end_year <= 2020,
         country == "India") %>%
  mutate(year = (year %/% 5) * 5,
         city = fct_lump(city, 5)) %>% 
  
  ggplot(aes(year, cost_km_millions, group = year)) +
  # geom_boxplot(outlier.size = -1) +
  geom_violin(outlier.size = -1, draw_quantiles = c(.25,.5,.75)) +
  geom_sina(aes(color = city)) +
  # geom_jitter(aes(color = city), height = 0, width = 1) +
  expand_limits(y = 0) +
  labs(y = "Cost/km (Real USD in Millions)",
       x = "Year",
       title = "Cost distr./ km in India")
```

```{r}
transit_cost %>%
  filter(
    # tunnel_per == 1,
    # end_year <= 2020,
         country == "India") %>%
  # mutate(year = (year %/% 5) * 5,
  #        city = fct_lump(city, 5)) %>% 
  
  ggplot(aes(x = city, y = cost_km_millions, color = year)) +
  # geom_boxplot(outlier.size = -1) +
  geom_violin(outlier.size = -1, draw_quantiles = c(.25,.5,.75)) +
  geom_sina() +
  # geom_jitter(aes(color = city), height = 0, width = 1) +
  expand_limits(y = 0) +
  labs(y = "Cost/km (Real USD in Millions)",
       x = "City",
       title = "Cost distr./ km in India") +
  scale_colour_viridis_b()
```


```{r}
transit_cost %>%
  filter(tunnel_per == 1,
         end_year <= 2020,
         country == "China") %>%
  mutate(city = fct_lump(city, 4)) %>%
  
  ggplot(aes(x = stations / length, y = cost_km_millions, 
             size = length, color = city)) +
  geom_point() +
  expand_limits(x = 0, y = 0) +
  labs(x = "Stations / km", "Cost / kilometer",
       y = "Cost / km")
```

```{r}
transit_cost %>%
  filter(
    # tunnel_per == 1,
    # end_year <= 2020,
    country == "India") %>%
  mutate(city = fct_lump(city, 4)) %>%
  
  ggplot(aes(x = length/ stations , y = cost_km_millions, 
             size = length, color = city)) +
  geom_point() +
  expand_limits(x = 0, y = 0) +
  labs(title = "India: Avg Distance b/w Stations Vs Cost",
       subtitle = "Size of bubble depends on length of Project",
       x = "Distance b/w Stations (km)", y = "Cost / kilometer")
```


```{r}
transit_cost %>%
  # filter(
    # tunnel_per == 1,
    # end_year <= 2020,
    # country == "India") %>%
  mutate(country = fct_lump(country, 15)) %>%
  
  ggplot(aes(x = length/ stations , y = cost_km_millions, 
             size = length, color = country)) +
  geom_point() +
  expand_limits(x = 0, y = 0) +
  labs(title = "World: Avg Distance b/w Stations Vs Cost",
       subtitle = "Size of bubble depends on length of Project",
       x = "Distance b/w Stations (km)", y = "Cost / kilometer")
```














































